---
title: "p8105_hw3_al4225"
author: "Anjing"
date: "2022-10-10"
output: github_document
---

```{r}
library(tidyverse)
```

## Problem 2

### Wrangling
Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

There are `r accel_data = read_csv("./data/accel_data.csv",
                col_names = TRUE) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(
    day_type = ifelse(day %in% c("Saturday", "Sunday"), "weekday", "weekend")) %>%
  group_by(week,day_id) %>%
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity_minute", 
    values_to = "activity_counts",
    names_prefix = "activity_") %>%
  mutate(
    activity_counts = round(activity_counts, digits = 0),
    activity_minute = as.integer(activity_minute)) %>%
    select(week, day_id, day, day_type, everything())
nrow(accel_data)` observations and `r accel_data = read_csv("./data/accel_data.csv",
                col_names = TRUE) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(
    day_type = ifelse(day %in% c("Saturday", "Sunday"), "weekday", "weekend")) %>%
  group_by(week,day_id) %>%
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity_minute", 
    values_to = "activity_counts",
    names_prefix = "activity_") %>%
  mutate(
    activity_counts = round(activity_counts, digits = 0),
    activity_minute = as.integer(activity_minute)) %>%
    select(week, day_id, day, day_type, everything())
ncol(accel_data)` variables. Each observation represents the activity counts for each minute of a 24-hour day starting at midnight in the day of given week. There are some key variables such as week, day_id, day, day_type and activity_minute and activity_counts. 
```{r}
accel_data = read_csv("./data/accel_data.csv",
                col_names = TRUE) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(
    day_type = ifelse(day %in% c("Saturday", "Sunday"), "weekday", "weekend")) %>%
  group_by(week,day_id) %>%
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity_minute", 
    values_to = "activity_counts",
    names_prefix = "activity_") %>%
  mutate(
    activity_counts = round(activity_counts, digits = 0),
    activity_minute = as.integer(activity_minute)) %>%
    select(week, day_id, day, day_type, everything())
nrow(accel_data)
ncol(accel_data)

accel_data
```

### Table for total activity
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

According to the table, we can see that there are 35 total activity counts in 35 days. From week1 to week3, the total activity counts maintain a high level and gradually increase from week1 to week2. From week3 to week5, the total activity counts tend to decrease especially on Saturday. 
```{r}
accel_data %>%
  group_by(week, day_id, day) %>%
  summarize(activity_total = sum(activity_counts)) %>%
  knitr::kable()
accel_data
```

### Graph
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

Description:
1.There are two evident peaks in the plot which are at 21:00 on Friday and during 9:00 to 12:00 on Sunday.  
2.The activity counts are higher on Friday and Sunday so this 63 year-old male with BMI 25 prefers to do the activity on Friday and Sunday.
3.The activity counts decline from 21:00 to 3:00 the next day and they increase from 3:00 to 21:00. It means that the man tends to have a rest after 21:00.  
```{r}
accel_data %>%
  ggplot(aes(x = activity_minute, y = activity_counts, color = day)) + 
  geom_smooth(se = FALSE) +
  labs(
    title = "activity_counts plot",
    x = "time of the day",
    y = "activity counts"
  ) +
  scale_x_continuous(
    breaks = c(0, 360, 720, 1080, 1440), 
    labels = c("0:00", "6:00", "12:00","18:00", "24:00" )
  ) +
  theme(legend.position = "bottom")
  
```


## Problem 3
